PART ONE
Loading and preprocessing the data
```{r}
setwd("C:/Users/Pratul/Downloads/repdata_data_activity/")
activity <- read.csv("C:/Users/Pratul/Downloads/repdata_data_activity/activity.csv",colClasses = c("numeric", "character","integer"))
```
PART TWO
Process/transform the data
```{r}
dim(activity)
head(activity)
tail(activity)
summary(activity)
names(activity)
str(activity)
library(plyr)
library(dplyr)
library(lubridate)
library(ggplot2)
total.steps <- tapply(activity$steps, activity$date, FUN = sum, na.rm = TRUE)
activity$date <- ymd(activity$date)
```
PART THREE
What is mean total number of steps taken per day?
```{r}
mean(total.steps)
median(total.steps)
```


Calculate the total number of steps taken per day
```{r}
steps <- activity %>%
  filter(!is.na(steps)) %>%
  group_by(date) %>%
  summarize(steps = sum(steps)) %>%
  print  
```


Make a histogram of the total number of steps taken each day 
```{r, echo=FALSE}
ggplot(steps, aes(x=date, y=steps))+geom_histogram(stat="identity")+ xlab("Dates")+ ylab("Steps")+ labs(title= "Total numbers of Steps per day")
```

PART FOUR
What is the average daily activity pattern?
```{r}
daily <- activity %>%
        filter(!is.na(steps)) %>%
        group_by(interval) %>%
        summarize(steps=mean(steps)) %>%
        print
```

Make a time series plot (i.e. type = ?l?) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r, echo=FALSE}
plot(daily, type = "l")
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
daily[which.max(daily$steps), ]$interval
```

PART FIVE
Imputing missing values
```{r}
missing <- sum(is.na(activity))
```

Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
new <- activity %>%
        group_by(interval) %>%
        mutate(steps = ifelse(is.na(steps), mean(steps, na.rm=TRUE), steps))
summary(new)
```

Make a histogram of the total number of steps taken each day
```{r}
new.steps <- new %>%
  group_by(date) %>%
  summarize(steps = sum(steps)) %>%
  print  
```

```{r, echo=FALSE}
ggplot(new.steps, aes(x=date, y=steps))+geom_histogram(stat="identity")+ xlab("Dates")+ ylab("Imputed Steps")+ labs(title= "Total numbers of Steps per day (missing data imputed)")
```

Calculate and report the mean and median total number of steps taken per day.
```{r}
imputed.steps <- tapply(new$steps, new$date, FUN = sum, na.rm = TRUE)
new$date <- ymd(new$date)
mean(imputed.steps)
```

Do these values differ from the estimates from the first part of the assignment?
```{r}
mean(total.steps)==mean(imputed.steps)
median(total.steps)==median(imputed.steps)
summary(total.steps)
summary(imputed.steps)
```

What is the impact of imputing missing data on the estimates of the total daily number of steps?
```{r}
summary(imputed.steps) - summary(total.steps)
```

PART SIX
Are there differences in activity patterns between weekdays and weekends?
```{r}
dayofweek <- function(date) {
    if (weekdays(as.Date(date)) %in% c("Saturday", "Sunday")) {
        "weekend"
    } else {
        "weekday"
    }
}
new$daytype <- as.factor(sapply(new$date, dayofweek))
```

Make a panel plot containing a time series plot
```{r, echo=FALSE}
par(mfrow = c(2, 1))
for (type in c("weekend", "weekday")) {
    steps.type <- aggregate(steps ~ interval, data = new, subset = new$daytype == 
        type, FUN = mean)
    plot(steps.type, type = "l", main = type)
}
```

```{r}
sessionInfo()
```

